---
title: "R Notebook"
output: html_notebook
---
```{r}

# all packages used for the assignment
library(mosaic)
library(tidyverse)
library(mosaicData)
library(DataComputing)
library(ggplot2)
library(data.table)

```

```{r}
VIX<-read.table("VIX.csv", header=TRUE,sep=",",stringsAsFactors = FALSE)
GLD<-read.table("GLD.csv", header=TRUE,sep=",",stringsAsFactors = FALSE)
USO<-read.table("USO.csv", header=TRUE,sep=",",stringsAsFactors = FALSE)
FL<-read.table("FL interest etf.csv", header=TRUE,sep=",",stringsAsFactors = FALSE)
StandardPoor <-read.table("S&P.csv", header=TRUE,sep=",",stringsAsFactors = FALSE)
```

# Guiding Question (i.e., Research Question): Make a clear statement of a single guiding/research question that will drive your analysis.  It should be phrased as a question that you will try to investigate (i.e., it should have a question mark) 
## The guiding question for our group is how to different stock indexes and prices of various asset classes relate to each other in price, and are there correlations between them?

# Become acquainted with your data sources:
## Where did you find them?
### Yahoo Finance
## Who collected/maintains them?
### Yahoo Finance collects and maintains the data. 
## When & Why were they originally collected?
### This data is collected continually by Yahoo Finance to analyze stock prices.
## What does a case represent in each data source, and how many total cases are available?
### Each case represents a stock on a given day, and each data set has 504 cases, spanning 2 years of trading.
## What are some of the variables that you plan to use
### We plan to use the closing price and date variables.

#The R Notebook should include R code chunks & output as well as narrative (bullet list is fine at this stage) as you answer these questions
## Explore intuition related to the research question
### The preliminary idea is that each asset class will trade in a specific way to the overall market (which we chose to be the S&P 500, further details regarding each asset/stock will be included in the final draft).  Lower interest rates stimulate the economy, gold prices are generally considered a safe-haven asset, and VIX measures volatility through S&P500 options, thus all three should be fairly inverse to the S&P 500. Finally, we are unsure about oil be we theorize that it will be positively correlated to the S&P as oil demand will likely follow the rest of the market.
## Create some informative plots and summary statistics 
```{r}
MasterData <-
  StandardPoor %>%
  select(Date, spClose = Close) %>%
  left_join(FL %>% select(Date, flClose = Close), by = "Date")%>%
  left_join(GLD %>% select(Date, gldClose = Close), by = "Date")%>%
  left_join(USO %>% select(Date, usoClose = Close), by = "Date")%>%
  left_join(VIX %>% select(Date,vixClose = Close), by = "Date")
  
```
```{r}
StandardPoor %>%
  ggplot(aes(x=Date, y=Close))+geom_line(aes(y=Close)
```


# TWO Data Sources (at least)
## Primary data:  There are 5 primary data sources from price data on 5 various indexes of assets.
